### 2015-02-07 src/paper4/p4_discussion.tex
Updated paper

We investigated whether the eye movements used in self-motion perception \cite{clemens2015a} are properly scaled by fixation distance, as required by the geometry. Experiments assessed self-motion perception during either body- and world-centered fixation at two different fixation depths. During far world-centered fixation (where eye movements are small) self-motion was underestimated compared to nearby world-centered fixation (where eye movements are large). Fixation depth did not influence self-motion perception during body-centered fixation (where eye movements are absent). To quantify the relative depth-dependent scaling of eye movements for nearby and far away fixation targets, we fitted a simple linear model to the perceptual and oculomotor responses across four conditions. While two participants show partial scaling, the six remaining participant did not show any sign of scaling.

In addition to the absence of scaling, another explanation is that eye movement information is integrated in a statistically optimal fashion, taking signal noise into account. In the world-fixed condition, this would mean that noisy eye movements in the far fixation are weighed less than those in the less noisy ones in the near fixation condition, giving rise to the observed difference. In this case, similar optimal integration should also occur in the body-fixed fixation condition. As no such differences between the near and far body-fixed fixation conditions have been observed, this is unlikely to be the case.

As it becomes impossible to distinguish between body- and world-fixed fixation point at far fixation distances, the lack of scaling could - in theory - be explained by participants incorrectly perceiving both the body- and world-fixed far fixation points as being body-fixed. This is unlikely because the target displacement associated with a world-fixed target was between X and Y degrees in our experiment, which is easily perceivable especially given that the target was foveated. This adds confidence to our hypothesis that eye movements indeed influence self-motion perception, with moderate to no scaling for fixation depth.

% Relation to Clemens, 2015
\begin{figure}
    \includegraphics[width=0.5\textwidth]{src/paper4/paper4_figure6.eps}

@@

[Conclusion]

>>>>

% Relation to previous work / Should rewrite this paragraph!
In our previous study, we demonstrated that oculomotor signals play a substantial role in the perception of self-motion, even in the absence of optic flow or any other visual stimulation \cite{clemens2015a}. Although the vestibular system provided the most significant contribution, oculomotor signals were shown to account for about 20\% of the overall percept. Because these experiments were performed with a single fixation depth, it was not clear whether the brain weighted the oculomotor signal in a depth-dependent manner when using it as a translation cue, or merely uses the signal as a rudimentary cue to self-motion. In the present study we tested between these two possibilities.

% Basic observations
We assessed self-motion perception during either body- and world-centered fixation at two different fixation depths. Our results show that self-motion was underestimated when comparing far and near fixation trials in the world-centered condition, which argues against a proper scaling of the eye movement signal. Fixation depth did not influence self-motion perception during body-centered fixation (where eye movements are virtually absent).

% Model results
To quantify the relative depth-dependent scaling of eye movements for nearby and far away fixation targets, we fitted a straightforward linear model to the perceptual responses based on the oculomotor behavior across four conditions. While two participants show partial scaling, the other six participants did not show any sign of scaling. Thus, we conclude that while oculomotor signals provide a robust clue to translation perception, they are not properly scaled by fixation depth.

\subsection{Relation to previous studies}

% Relation to Clemens, 2015
\begin{figure}
    \includegraphics[width=0.5\textwidth]{src/paper4/p4_figure6.pdf}

@@

% Talk about whether perception and action employ different mechanisms?
% Check work by Merfeld
%
% Vestibular perception and action employ qualitatively different mechanisms I and II
% Merfeld, Park, Gianna-Pulin, Black, Wood, 2005
%

% -- %

% Self-motion perception; disambiguation
% Rotation perception; do not expect fixation distance to play a role

% Link to path reception during rotation, influence of instructions, depth range, and dot density:
% Li Li, William H. Warren Jr. doi:10.1016/j.visres/2004.03.008

% Experimental Brain Research doi:10.1007/s00221-009-1828-z 123
% Eye eccentricity modifies the perception of whole-body rotation
% Gaelle Quarck, Lena Lhuisset, Olivier Etart, Pierre Denise

\subsection{Alternative explanations}

In addition to the absence of scaling, another explanation is that eye movement information is integrated in a statistically optimal fashion, taking signal noise into account. In the world-fixed condition, this would mean that noisy eye movements in the far fixation are weighed less than those in the less noisy ones in the near fixation condition, giving rise to the observed difference. In this case, similar optimal integration should also occur in the body-fixed fixation condition. As no such differences between the near and far body-fixed fixation conditions have been observed, this is unlikely to be the case.  \footnote{\textbf{I do not understand the text Pieter proposed, I think it says something completely different.} While the brain should scale eye movements by fixation depth when using them as a geometrical veridical cue to translation perception, our data do not confirm this hypothesis. Does this imply that the brain does not apply any transformation to oculomotor feedback signals? It is important to realize that our analysis did not account for the fact that noise levels in the oculomotor signal could depend on fixation depth. If we would incorporate this notion in our analysis, i.e.  oculomotor signals are weighted by their noise level, it would suggest that the eye movements occurring with near fixation points are more noisy than those with a far fixation point. Our data provide no indication for this being the case. }

% FIXME: Change X and Y in paragraph below; I don't like asking questions like this.
Could the lack of scaling be explained by how participants perceive the far fixation point? Because the difference between body- and world-fixed fixation points is reduced at far fixation distances, the lack of scaling could - in theory - be explained by participants incorrectly perceiving both the body- and world-fixed far fixation points as being body-fixed. We consider this an unlikely explanation, because the target displacement associated with a world-fixed target was between X and Y degrees in our experiment, which is easily perceivable especially given that the target was foveated. This adds confidence to our claim that eye movements indeed influence self-motion perception, with moderate to no scaling for fixation depth.

\subsection{Conclusion}
% Conclusion
In summary, fixation depth is not properly taken into account, suggesting that a relatively raw eye movement signal is used in self-motion perception. While efference copies of the eye movement signals or proprioceptive feedback could be directly responsible for the perceived self-motion, it cannot be ruled out that the signals that drive the eye movements are not causing the reported effects instead.

### 2015-02-07 src/paper4/p4_intro.tex
Updated paper

An accurate estimate of self-motion is important when navigating through our environment. During navigation, we typically keep track of the objects we interact with or do not want to bump into using eye movements. The shift of the retinal image caused by self-motion (i.e., optic flow) is disturbed by these eye movements. However, using the oculomotor signal, the brain accounts for these disturbances by internally separating optic flow in two components, one caused by self-motion and the other by eye movement (references).

@@

In the present study, we investigate how fixation distance influences perception of self-motion across passive lateral translation. Using a psychophysical approach, participants had to indicate whether the second body displacement of two one-second translation intervals was smaller or longer than the first. We show that translation amplitude is perceived smaller when fixating a far rather than a nearby world-centered target, indicating that eye movements are not properly scaled for self-motion perception. Together with the further observation that self-motion perception with body-centered fixation targets is not affected by fixation depth, we conclude that eye movements are merely a rudimentary cue to self-motion, with a compensation for fixation depth that is partial at best.

>>>>

% Changing estimate this to control sounds weird as we investigate passive self-motion.
An accurate estimate of self-motion is important when navigating through our environment. During navigation, we typically keep track of the objects we interact with or do not want to bump into using eye movements. The shift of the retinal image caused by self-motion (i.e., optic flow) is disturbed by these eye movements. However, using the oculomotor signal, the brain accounts for these disturbances by internally separating optic flow in two components, one caused by self-motion and the other by eye movement (references).

@@

In the present study, we therefore investigate how fixation distance influences perception of self-motion across passive lateral translation. Using a psychophysical approach, participants had to indicate whether the second body displacement of two one-second translation intervals was smaller or longer than the first. We show that translation amplitude is perceived smaller when fixating a far compared to a nearby world-centered target, indicating that eye movements are not properly scaled in self-motion perception. Together with the further observation that self-motion perception with body-centered fixation targets is not affected by fixation depth, we conclude that eye movements are merely a rudimentary cue to self-motion, with a compensation for fixation depth that is partial at best.

### 2015-02-07 src/paper4/p4_methods.tex
Updated papers
Updated paper

We used a two-alternative forced choice (2-AFC) task to study the influence of fixation depth on the perception of linear translation. We tested two fixation depths: near (50 cm) and far (200 \si{\centi\metre}) and two different eye fixation conditions: world-, and body-fixed fixation. A trial contained two sequential motion intervals of equal duration (1s), with the motion in the same direction (either leftward or rightward). Fixation condition (world-, or body-fixed) was the same in both intervals, but a different fixation depth was presented in each. Participants had to judge whether the translation during the second motion interval was longer or shorter than in the first motion interval.

\begin{figure}
    \includegraphics[width=1.0\textwidth]{src/paper4/paper4_figure1.eps}

@@

The timing of a single trial is shown in \figref{p4:fig1}A. Each trial started with the onset of a central fixation point (i.e. aligned between the eyes) for 0.5s. Subsequently, the first 1s motion interval commenced, during which the fixation type which was at on the the two depths, remained stationary (world) or moved along with the participant (body). Subsequently a 1.75s delay followed in which the participant was kept in complete darkness. Then, the central fixation point reappeared, followed 0.5 \si{\second} later by a second 1 \si{\second} translation interval, with the same fixation condition but with the fixation point at the other depth as in the first interval. Finally, the participant responded by moving a 1-dimensional joystick away from (longer) or towards (shorter) the body. Top-view illustrations of body-centered and world-centered trials are shown in \figref{p4:fig1} panels B and C.

@@

For each combination of the two main conditions and the two reference/probe orders (see \tabref{p4:tab1}), we quantified the perceived probe translation by calculating the probability the probe translaton was perceived longer than the reference translation as a function of actual probe translation. We used a maximum likelihood fit of a cumulative Gaussian function to summarize the psychometric data:

@@

Using a straightforward model, we investigate to what extent fixation is taken into account in the contribution of eye movements to self-motion perception. As in \citeA{clemens2015a}, we model the perceived translation, $p_i$, as a weighted combination of a vestibular, $m_i$, and oculomotor estimate of translation, $\hat{\varphi}_i$ (\eqnref{p4:eq4}).

@@

% This seems a bit flaky!
Variable $i$ represents either the reference, $r$, or probe, $p$, interval. Ideally, to server a veridical self-motion cue, the brain must scale the eye movement by the depth of fixation, $d$. This scaling is reflected by parameter, $\alpha_{d_i}$. If this parameter is the same across the two fixation depths, $d$, then there is not depth modulation of the oculomotor estimate of translation.

By definition, at the PSE, the probe translation is perceived as equal in length to the 10 cm reference displacement, $p_r = p_p$. By substituting both sides by the right hand side of \eqnref{p4:eq4}, we then obtain:

@@

We fit \eqnref{p4:eq5} to the data using linear regression, finding one weight for every fixation depth (that is, $\alpha_{50}$ and $\alpha_{200}$) that minimizes the sum of squared errors ($\Sigma \epsilon^2$). Because these parameters can, in theory, contain both a depth dependent and a depth independent scaling, we compute their ratio, $\frac{\alpha_{50}}{\alpha_{200}}$, to remove any depth independent components. In case of perfect compensation, the expected ratio is $\frac{50}{200} = 1/4$, while in case of no depth-dependent scaling it is 1.

>>>>

We used a two-alternative forced choice (2-AFC) task to study the influence of fixation depth on the perception of linear translation. We tested two fixation depths: near (50 cm) and far (200 \si{\centi\metre}) and two different eye fixation conditions: world-, and body-fixed fixation. A trial contained two sequential motion intervals of equal duration (1s), with the motion in the same direction (either leftward or rightward).

%Fixation condition (world-, or body-fixed) was the same in both intervals, but a different fixation depth was presented in each. Participants had to judge whether the translation during the second motion interval was longer or shorter than in the first motion interval.

\begin{figure}
    \includegraphics[width=1.0\textwidth]{src/paper4/p4_figure1.pdf}

@@

The timing of a single trial is shown in \figref{p4:fig1}A. Each trial started with the onset of a central fixation point (i.e. aligned between the eyes) for 0.5s. Subsequently, the first 1s motion interval commenced, during which the fixation type which was at on the two depths, remained stationary (world condition) or moved along with the participant (body condition). Subsequently a 1.75s delay followed in which the participant was kept in stationary complete darkness. Then, the central fixation point reappeared, followed 0.5 \si{\second} later by a second 1 \si{\second} translation interval, with the same fixation condition but with the fixation point at the other depth as in the first interval. Finally, the participant responded by moving a 1-dimensional joystick away from (longer) or towards (shorter) the body. Top-view illustrations of body-centered and world-centered trials are shown in \figref{p4:fig1} panels B and C.

@@

For each combination of the two main conditions and the two reference/probe orders (see \tabref{p4:tab1}), we quantified the perceived probe translation by calculating the probability the probe translation was perceived longer than the reference translation as a function of actual probe translation. We used a maximum likelihood fit of a cumulative Gaussian function to summarize the psychometric data:

@@

Using a straightforward model, we investigate to what extent fixation depth is taken into account in the contribution of eye movements to self-motion perception. As in \citeA{clemens2015a}, we model the perceived translation, $p_i$, as a weighted combination of a vestibular, $m_i$, and oculomotor estimate of translation, $\hat{\varphi}_i$ (\eqnref{p4:eq4}).

@@

% This seems a bit flaky!
Variable $i$ represents either the reference, $r$, or probe, $p$, interval. Ideally, to server a veridical self-motion cue, the brain must scale the eye movement by the depth of fixation, $d$. This scaling is reflected by parameter, $\alpha_{d_i}$. If this parameter is the same across the two fixation depths, $d$, then there is no depth-dependent modulation of the oculomotor estimate of translation.

By definition, at the PSE, the probe translation is perceived as equal in length to the 10 cm reference translation, $p_r = p_p$. By substituting both sides by the right hand side of \eqnref{p4:eq4}, we then obtain:

@@

We fit \eqnref{p4:eq5} to the data using linear regression, finding one weight for every fixation depth (that is, $\alpha_{50}$ and $\alpha_{200}$) that minimizes the sum of squared errors ($\Sigma \epsilon^2$). Because these parameters can, in theory, contain both a depth-dependent and a depth-independent scaling, we compute their ratio, $\frac{\alpha_{50}}{\alpha_{200}}$, to remove any depth independent components. In case of perfect compensation, the expected ratio is $\frac{50}{200} = 1/4$, while in case of no depth-dependent scaling it is 1.

### 2015-02-07 src/paper4/p4_results.tex
Updated paper 4
Updated paper

In a recent study we have shown that eye movements serve as a cue in the perception of body translation, even in the absence of optic flow \cite{clemens2015a}. Because eye rotations must be weighed by target depth ($\varphi d = T$) to serve as an adequate translation cue, we tested the effect of near versus far fixation on self-motion perception. Participants were presented with two subsequent translation movements \figref{p4:fig1} and had to judge whether the second translation was perceived longer or shorter than the first. During each interval, participants fixated either a body- or world-fixed fixation point either nearby (50 cm) or far away (200 cm) from the participant.

@@

We first tested the ability of participants to follow the eye movement instruction in our paradigm. \figref{p4:fig2}A depicts exemplar eye traces for the 10cm reference translation with nearby and far fixation points in both the body and world condition. Changes in gaze are largely absent in the body near and body far conditions (brown and orange respectively), as required. During the world conditions, the eye movements were large when fixating nearby targets (purple) and small when fixating far away ones (pink), which reflects the geometrical constraints.

We compressed the eye movement data by taking the average ratio between the measured eye excursion and the geometrically required displacement were the target at a world-fixed location. \figref{p4:fig2}B shows these normalized eye displacements for all participants, which is about zero during body fixed fixations (brown and orange) and close to one in the world fixed fixations (purple and pink).

@@

	\caption{\panelref{A} and \panelref{B} Psychometric curves (colored lines) and associated binned data (circles) for one participant. Circle size represents the amount of trials within the bin. Psychometric curves before collapsing across reference order are shown as gray lines. \panelref{A} Body-fixed condition (brown) while fixation was either near (dark) or far (light). \panelref{B} World-fixed condition (purple) while fixation was either near (dark) or far (light).
	\panelref{C} and \panelref{D} Individual and average points of subjective equality (PSEs). Color scheme matches panels A. and B.
	}
	\label{p4:fig3}
\end{figure}

To answer this question, \figref{p4:fig3} illustrates psychophysical data on self-motion perception of a single participant for the two-fixation depths in the body (\figref{p4:fig3}A) or world condition (\figref{p4:fig3}B) respectively. Lighter and darker colors indicate which fixation type was the reference translation (see Legend). The influence of fixation distance is characterized by a shift of the psychometric functions relative to the 10 \si{\centi\metre} reference translation (i.e. the PSE). For example, the rightward shift of the pink curve in \figref{p4:fig3}B, representing the world-condition, means that with a far target a longer translation (~15 \si{\centi\metre}) was required to be perceived equivalent to a 10 \si{\centi\metre} reference translation with nearby fixation. Likewise, the leftward shift of the purple curve indicates that a shorter translation with near fixation (~6 \si{\centi\metre}) is required to be perceived the same as the 10 \si{\centi\metre} reference translation with far fixation. Together, these opposite biases suggest that translations are perceived shorter for fixations further away. The brown and orange curves in \figref{p4:fig3}A do not show such shifts, indicating that fixation depth (i.e. near versus far) has no effect in absence of eye movements (i.e. when fixating a body-fixed target during the translation).

@@

\begin{wrapfigure}{l}[5pt]{0.5\textwidth}
    \includegraphics[width=1.0\textwidth]{src/paper4/p4_figure4.pdf}

	\caption{PSEs for all participants and the average \textpm SE. \panelref{A}  Body-fixed condition while fixation was either near (brown) or far (orange). \panelref{B} World-fixed condition while fixation was either near (purple) or far (pink).}
	\label{p4:fig4}
\end{wrapfigure}

In order to quantify how much fixation depth scales the eye movement as a linear self-motion cue, we used a simple linear model (see \nameref{p4:sec:methods}). This model explains the perceived translation distance as a weighted average of a vestibular estimate, equal to the actual translation, and an oculomotor-based estimate. 
We used one weighting parameter, $\alpha$, per fixation depth, $\alpha_{50}$ and $\alpha_{200}$ (see \tabref{p4:tab2} for best-fit values). Using these parameters we predicted the PSEs, i.e. $m_p$ in \eqnref{p4:eq5}, and plotted it against the actually  observed PSEs in \figref{p4:fig4}. The positive correlation ($\rho = 0.xx$, $p < 0.01$) between observed and predicted PSEs shows that our simple model does reasonably well in predicting perceptual performance.

\begin{wrapfigure}{l}[5pt]{0.5\textwidth}
    \includegraphics[width=1.0\textwidth]{src/paper4/p4_figure5.pdf}

	\caption{}
	\label{p4:fig5}
\end{wrapfigure}

By taking the ratio of these weighting parameters, any depth independent effects were removed. As this ratio contains only depth dependent effects, it is equal to the ratio of depths used for eye movement scaling. In case there is no depth scaling, i.e. when $d_{50} = d_{200}$, the expected ratio is one. Perfect compensation, that is when $d_{50} = 50  \wedge d_{200} = 200$, the expected ratio is $1/4$. The actual ratio between $d_{50}$ and $d_{20}$ is plotted for each participant in \figref{p4:fig5}. While two participants show moderate compensation for depth, the majority of participants show no sign of scaling of eye movements by fixation distance. This is consistent with the observation in the world condition that translations are perceived shorter with far compared to near fixations.

>>>>

In a recent study we have shown that eye movement signals contribute to the perception of body translation, even in the absence of optic flow \cite{clemens2015a}. Because eye rotations must be scaled by target depth ($\varphi d = T$) to serve as an adequate translation cue, we tested the effect of near versus far fixation on self-motion perception. Participants were presented with two subsequent translation movements \figref{p4:fig1} and had to judge whether the second translation was perceived longer or shorter than the first. During each interval, participants fixated either a body- or world-fixed fixation point either nearby (50 cm) or further away (200 cm).

@@

We first tested the ability of participants to follow the eye movement instruction in our paradigm. \figref{p4:fig2}A depicts exemplar eye traces for the 10cm reference translation with nearby and far fixation points in both the body and world condition. Changes in gaze are largely absent in the body near and body far conditions (brown and orange traces respectively), as required. During the world conditions, the eye movements were large when fixating nearby targets and small when fixating far away ones (purple and pink traces respectively), which reflects the geometrical constraints.

We analyzed the eye movement data by taking the average ratio between the measured eye excursion and the geometrically required displacement were the target at a world-fixed location. \figref{p4:fig2}B shows these normalized eye displacements for all participants, which are about zero during body fixed fixations (brown and orange) and close to one in the world fixed fixations (purple and pink data points). Does the brain scale these eye movements by fixation depth in order to interpret them as a linear self-motion cue?

@@

	\caption{\panelref{A} and \panelref{B} Psychometric curves (colored lines) and associated binned data (circles) for one participant. Circle size represents the amount of trials within the bin. Psychometric curves before collapsing across reference order are shown as gray lines. \panelref{A} Body-fixed condition (brown) while fixation was either near (dark) or far (light). \panelref{B} World-fixed condition (purple) while fixation was either near (dark) or far (light).
	\panelref{C} and \panelref{D} Individual and average points of subjective equality (PSEs). Color scheme matches panels A and B.
	}
	\label{p4:fig3}
\end{figure}

To answer this question, \figref{p4:fig3} illustrates psychophysical data on self-motion perception of a single participant for the two-fixation depths in the body (\figref{p4:fig3}A) and world condition (\figref{p4:fig3}B) respectively. Lighter and darker colors indicate which fixation type was the reference translation (see figure legend). The influence of fixation distance is characterized by a shift of the psychometric functions relative to the 10 \si{\centi\metre} reference translation (i.e. the PSE). For example, the rightward shift of the pink curve in \figref{p4:fig3}B, representing the world-condition, means that with a far target a longer translation (~15 \si{\centi\metre}) was required to be perceived equivalent to a 10 \si{\centi\metre} reference translation with nearby fixation. Likewise, the leftward shift of the purple curve indicates that a shorter translation with near fixation (~6 \si{\centi\metre}) is required to be perceived the same as the 10 \si{\centi\metre} reference translation with far fixation. Together, these opposite biases suggest that translations are perceived shorter for fixations further away. The brown and orange curves in \figref{p4:fig3}A do not show such shifts, indicating that fixation depth (i.e. near versus far) has no effect in absence of eye movements (i.e. when fixating a body-fixed target during the translation).

@@

\begin{figure}
    \includegraphics[width=0.5\textwidth]{src/paper4/p4_figure4.pdf}

	\caption{PSEs for all participants and the average \textpm SE. \panelref{A}  Body-fixed condition while fixation was either near (brown) or far (orange). \panelref{B} World-fixed condition while fixation was either near (purple) or far (pink).}
	\label{p4:fig4}
\end{figure}

In order to quantify how much fixation depth scales the eye movement as a linear self-motion cue, we used a simple linear model (see \nameref{p4:sec:methods}). This model explains the perceived translation distance as a weighted average of a vestibular estimate, equal to the actual translation, and an oculomotor-based estimate. 
We used two weighting parameters, $\alpha_{50}$ and $\alpha_{200}$, one per fixation depth (see \tabref{p4:tab2} for best-fit values). Using these parameters we predicted the PSEs, i.e. $m_p$ in \eqnref{p4:eq5}, and plotted them against the actually observed PSEs in \figref{p4:fig4}. The positive correlation ($\rho = 0.xx$, $p < 0.01$) between observed and predicted PSEs shows that our simple model does reasonably well in predicting perceptual performance.

\begin{figure}
    \includegraphics[width=0.5\textwidth]{src/paper4/p4_figure5.pdf}

	\caption{}
	\label{p4:fig5}
\end{figure}

By examining the ratio of these weighting parameters, we remove any depth-independent contributions. In absence of  depth scaling, i.e. when $d_{50} = d_{200}$, the ratio should be one. For perfect compensation, that is when $d_{50} = 50  \wedge d_{200} = 200$, the ratio should be $1/4$. The actual ratio between $d_{50}$ and $d_{200}$ is plotted for each participant in \figref{p4:fig5}. While two participants show moderate compensation for depth, the majority of participants show no sign of scaling of eye movements by fixation distance. This is consistent with the observation that translations are perceived shorter with far compared to near fixations in the world condition.
